```{r}
#install.packages("arrow")
```


```{r}
library(arrow)
library(tidyverse)

shd<- arrow::read_parquet("https://intro-datascience.s3.us-east-2.amazonaws.com/SC-data/static_house_info.parquet")

Temp<- arrow::read_parquet("https://intro-datascience.s3.us-east-2.amazonaws.com/SC-data/2023-houseData/102063.parquet")
head(Temp)
#ed <- arrow::read_parquet("energyData.parquet")

```


```{r}
wd <- read_csv("https://intro-datascience.s3.us-east-2.amazonaws.com/SC-data/weather/2023-weather-data/G4500010.csv")
```


```{r}
set.seed(1)
build100<-sample(shd$bldg_id,100,replace = FALSE)
build100
```


```{r}
library(tidyverse)
julyfilter<-function(data){
  filter<- data%>%filter(substr(as_datetime(time),6,7)=="07")
  return(filter)
}


```

```{r}
# Loop through each building ID
library(tidyverse)
mylist<- list()
for (i in build100) {
  url <- paste0("https://intro-datascience.s3.us-east-2.amazonaws.com/SC-data/2023-houseData/", i, ".parquet")
  
  # Read the Parquet file into a data frame
  df_name <- paste0("ed_", i)  # Create a dynamic variable name
  mylist<- append(mylist,df_name)
  assign(df_name, arrow::read_parquet(url))  # Assign the data frame to the dynamic variable
  df<-get(df_name)
  df$bldg_id<- i
  assign(df_name,df)
}

```


```{r}
df<-data.frame(as.character(mylist))
```

```{r}

for (i in df$as.character.mylist.) {
  filtered_name <- i
  assign(filtered_name, julyfilter(get(i)))
}
```

```{r}
fullEDf<- data.frame()

for (i in df$as.character.mylist.){
  fullEDf <- rbind.data.frame(fullEDf,get(i)) 
}

head(fullEDf)
summary(shd)
```

```{r}

merged_df<-merge.data.frame(shd,fullEDf,by = 'bldg_id', sort = TRUE, all = FALSE)
```


```{r}
nrow(unique.data.frame(merged_df))
```


```{r}
 install.packages("openxlsx")
```


```{r}
library(openxlsx)


# Specify the file path
excel_file <- "output.xlsx"

# Write the data frame to an Excel file
write.xlsx(merged_df, excel_file)

# Confirm the export
cat("Data frame exported to Excel file:", excel_file, "\n")
```


```{r}
#head(wd)
julywdincreased <- wd%>%filter(substr(as_datetime(date_time),6,7)=="07")%>%arrange(date_time)
julywdincreased$`Dry Bulb Temperature [°C]`<-julywdincreased$`Dry Bulb Temperature [°C]`+5.0

```


```{r}
sum(is.na(merged_df))
head(merged_df)
#name_column <- my_df[["Name"]]
# Explore unique values for categorical variables
unique_values <- lapply(merged_df, function(x) length(unique(x)))
unique_values_df <- data.frame(variable = names(unique_values), unique_count = unlist(unique_values))
head(unique_values_df)

cleaned_df <- merged_df[c('in.ahs_region','in.census_division','bldg_id','upgrade','weight','applicability','in.ashrae_iecc_climate_zone_2004','in.bedrooms','in.building_america_climate_zone','in.census_division','in.census_region','in.city','in.clothes_dryer','in.clothes_washer','in.clothes_washer_presence','in.county','in.dishwasher','in.ducts','in.electric_vehicle','in.emissions_electricity_units','in.federal_poverty_level','in.geometry_building_number_units_mf','in.geometry_building_type_acs','in.geometry_floor_area','in.geometry_foundation_type','in.geometry_wall_exterior_finish','in.geometry_wall_type','in.geometry_wall_type','in.holiday_lighting','in.hvac_cooling_efficiency','in.hvac_has_shared_system','in.hvac_heating_efficiency','in.hvac_shared_efficiencies','in.income','in.insulation_ceiling','in.insulation_floor','in.insulation_foundation_wall','in.insulation_rim_joist','in.insulation_roof','in.insulation_slab','in.insulation_wall','in.lighting','in.misc_extra_refrigerator','in.misc_freezer','in.misc_gas_fireplace','in.misc_gas_grill','in.misc_gas_lighting','in.misc_hot_tub_spa','in.misc_pool','in.misc_pool_heater','in.misc_pool_pump','in.misc_well_pump','in.natural_ventilation','in.occupants','in.plug_loads','in.refrigerator','in.sqft','in.state','in.tenure','in.usage_level','in.vacancy_status','in.vintage','in.water_heater_efficiency','in.water_heater_fuel','in.weather_file_city','in.weather_file_latitude','in.weather_file_longitude')]
head(cleaned_df)
summary(cleaned_df)
#bldg_id, in.sqft,in.ahs_region,wind speed,relative humidity,in.bedrooms,time,in.weather_file_latitude,in.weather_file_longitude,
summary(merged_df)
head(wd)
```


```{r}
head(merged_df)
tail(merged_df)

```


```{r}
unique_values2 <- lapply(shd, function(x) length(unique(x)))
unique_values_df2 <- data.frame(variable = names(unique_values2), unique_count = unlist(unique_values2))
head(unique_values_df2)
```


```{r}
unique_values3 <- lapply(fullEDf, function(x) length(unique(x)))
unique_values_df3 <- data.frame(variable = names(unique_values3), unique_count = unlist(unique_values3))
head(unique_values_df3)
```


```{r}
unique_values3 <- lapply(wd, function(x) length(unique(x)))
unique_values_df3 <- data.frame(variable = names(unique_values3), unique_count = unlist(unique_values3))
head(unique_values_df3)
```


```{r}
head(unique_values_df3)
```


```{r}
unique_values4 <- lapply(fullEDf, function(x) length(unique(x)))
unique_values_df4 <- data.frame(variable = names(unique_values4), unique_count = unlist(unique_values4))
head(unique_values_df4)
```


```{r}
str(merged_df)

```



```{r}
#_____________________________________________exploratory data analysis_____________________________

library(tidyverse) # For data manipulation and visualization
library(readxl) # For reading Excel files
library(ggplot2) # For advanced graphing
library(GGally) # For correlation plots
```



```{r}

```


```{r}
library(corrplot) # For another style of correlation matrix visualization
library(lubridate) # For date and time manipulation
#library(scales) # For scaling graph ele(ments
library(tidyverse)
library(caret) # For modeling preparation

data <- read_csv('C:/Users/nehac/Downloads/merged_shd_ed.csv-20231203T223336Z-001/merged_shd_ed.csv')
#head(Final_data)
energy_consumption_columns <- data %>%
  select(matches("^out\\..*consumption$"))

# Calculate the total energy consumption for each row
data$total_energy_consumption <- rowSums(energy_consumption_columns, na.rm = TRUE) 
#Creating new column that has the sum of the energy consumed by all the appliance
```

Are there any noticeable outliers in energy consumption that could indicate data entry errors or exceptional usage scenarios?

```{r}
# Select the energy consumption column (replace with the appropriate column name)
energy_consumption <- data$totalconsumption

# Create a box plot to visualize the distribution of energy consumption
ggplot(data, aes(x=in.county,y = energy_consumption)) +
  geom_boxplot(fill = "lightblue", color = "blue") +
  labs(
    y = "Energy Consumption",
    title = "Box Plot of Energy Consumption"
  ) +
  theme_minimal() +  coord_flip()

# Calculate summary statistics to identify outliers
summary_stats <- summary(energy_consumption)
Q1 <- quantile(energy_consumption, 0.25)
Q3 <- quantile(energy_consumption, 0.75)
IQR <- Q3 - Q1
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR

# Find outliers
outliers <- data %>%
  filter(energy_consumption < lower_bound | energy_consumption > upper_bound)

# Print summary statistics and outliers
cat("Summary Statistics:\n")
cat(paste("Min:", min(energy_consumption), "\n"))
cat(paste("1st Quartile (Q1):", Q1, "\n"))
cat(paste("Median (Q2):", median(energy_consumption), "\n"))
cat(paste("3rd Quartile (Q3):", Q3, "\n"))
cat(paste("Max:", max(energy_consumption), "\n"))
cat(paste("IQR (Interquartile Range):", IQR, "\n"))
cat(paste("Lower Bound:", lower_bound, "\n"))
cat(paste("Upper Bound:", upper_bound, "\n"))

cat("\nOutliers:\n")
print(outliers)
```

What is the Daily energy consumption accross the entire data set. 
```{r}
library(ggplot2)
library(dplyr)
library(lubridate)  # For date manipulation


# Convert the "time" column to a DateTime object
data$time <- as.POSIXct(data$time, format = "%d-%m-%Y %H:%M")

# Aggregate energy consumption by day
daily_energy_consumption <- data %>%
  group_by(Date = as.Date(floor_date(time, unit = "days"))) %>%
  summarize(DailyEnergyConsumption = sum(total_energy_consumption, na.rm = TRUE))

# Create a ggplot object for the daily energy consumption trend
ggplot(daily_energy_consumption, aes(x = Date, y = DailyEnergyConsumption)) +
  geom_line() +  # Create a line chart
  labs(
    x = "Date",
    y = "Daily Energy Consumption",
    title = "Daily Energy Consumption Trend"
  ) +
  theme_minimal()
```

What is the distribution of energy usage in July across residential properties in South Carolina and the relevant parts of North Carolina?

```{r}
# Filter the data for properties in South Carolina and relevant parts of North Carolina
filtered_data <- Final_data %>%
  filter(in.state %in% c("SC", "NC"))  # Replace "NC" with relevant parts of North Carolina

# Convert 'in.sqft' to numeric if it's not already
filtered_data$in.sqft <- as.numeric(filtered_data$in.sqft)

# Filter the data for the month of July
july_data <- filtered_data

# Create a factor variable for in.usage_level
july_data$in.usage_level <- factor(july_data$in.usage_level)

# Create a custom color palette
custom_colors <- c("green", "orange", "red")

# Create a histogram to visualize the distribution of energy usage in July
# Use the 'fill' aesthetic to specify custom colors
ggplot(july_data, aes(x = in.usage_level, fill = in.usage_level)) +
  geom_bar() +
  scale_fill_manual(values = custom_colors) +  # Specify custom colors
  labs(
    title = "Distribution of Energy Usage in July",
    x = "Energy Usage Level",
    y = "Frequency"
  )
```

```{r}

#Three Categories of Energy Usage: 
#The energy usage is categorized into three levels - High, Medium, and Low.

#Frequency of Each Category: 
#The vertical axis (y-axis) indicates the frequency of properties falling into each energy usage category.

#Most Common Usage Level: 
#The Medium energy usage category has the highest frequency, indicating that most properties consume a medium amount of energy.

#Potential Data Distribution: 
#The distribution seems to be unimodal with the peak at Medium usage, suggesting that energy usage among these residential properties is centered around a typical value with fewer properties showing extremely high or low usage.

```

Are there any noticeable trends or patterns in July based on the time of the day. 

```{r}


```
`
```{r}

```



How does energy usage in July vary across different counties within South Carolina ?
```{r}
# Calculate the total energy consumption by county
county_energy_consumption <- data %>%
  group_by(in.county) %>%
  summarize(TotalEnergyConsumption = sum(total_energy_consumption, na.rm = TRUE)) %>%
  arrange(desc(TotalEnergyConsumption))  # Sort by TotalEnergyConsumption in descending order

# Create a ggplot object to visualize county-wise energy consumption
ggplot(county_energy_consumption, aes(x = TotalEnergyConsumption, y = fct_reorder(in.county, desc(TotalEnergyConsumption)))) +
  geom_bar(stat = "identity", fill = "blue") +  # Create a horizontal bar plot
  labs(
    x = "Total Energy Consumption",
    y = "County",
    title = "County-wise Energy Consumption (Descending Order)"
  ) +
  theme_minimal() 
```

The buildings in the above counties that has maximum power consumption. 

```{r}
# Load required libraries
library(dplyr)


# Filter data for specific counties
counties_to_search <- c("G4500450", "G4500190", "G4500790", "G4500510", "G4500830", "G4500630", "G4500150")

filtered_data <- data %>%
  filter(in.county %in% counties_to_search)

# Find the building with maximum power consumption in each county
max_power_buildings <- filtered_data %>%
  group_by(in.county) %>%
  top_n(1, total_energy_consumption) %>%
  ungroup()

# Print the result
view(max_power_buildings)

```

Does the energy usage in July vary significantly depending on the weather conditions, such as temperature, humidity, or other meteorological factors?

```{r}


```

```{r}


```

What is the distribution of household characteristics such as the  income levels and how do these factors relate to July energy consumption?

```{r}
library(ggplot2)

# Create a grouped or stacked bar chart
custom_colors <- c("green", "orange", "red")
ggplot(data, aes(x = in.income, fill = total_energy_consumption)) +
  geom_bar(position = "stack", stat = "count") +
  scale_fill_manual(values = custom_colors)+
  labs(title = "Income vs Usage Levels",
       x = "Income Range",
       y = "Power consumption") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

****************************Feature Engineering****************************************

a. Correlation Analysis: Calculate the correlation between each feature and the target variable (total_energy_consumption). Features with higher absolute correlation values are more likely to affect the target variable.


```{r}

# Load required libraries
library(dplyr)
library(tidyr)

# Select relevant predictor variables (features) and the target variable (energy consumption)
predictor_variables <- data %>%
  select(bldg_id,
upgrade,
weight,
applicability,
in.sqft,
in.bedrooms,
in.building_america_climate_zone,
in.ceiling_fan,
in.city,
in.clothes_dryer,
in.clothes_washer,
in.clothes_washer_presence,
in.cooking_range,
in.cooling_setpoint,
in.corridor,
in.county,
in.dishwasher,
in.federal_poverty_level,
in.geometry_floor_area_bin,
in.geometry_foundation_type,
in.geometry_garage,
in.geometry_stories,
in.geometry_stories_low_rise,
in.heating_fuel,
in.heating_setpoint,
in.heating_setpoint_has_offset,
in.heating_setpoint_offset_magnitude,
in.hot_water_fixtures,
in.hvac_cooling_efficiency,
in.hvac_cooling_partial_space_conditioning,
in.hvac_cooling_type,
in.hvac_has_ducts,
in.hvac_has_zonal_electric_heating,
in.hvac_heating_efficiency,
in.hvac_heating_type,
in.hvac_heating_type_and_fuel,
in.income,
in.income_recs_2015,
in.income_recs_2020,
in.infiltration,
in.insulation_ceiling,
in.insulation_floor,
in.insulation_foundation_wall,
in.insulation_rim_joist,
in.insulation_roof,
in.insulation_slab,
in.insulation_wall,
in.lighting,
in.misc_extra_refrigerator,
in.misc_freezer,
in.misc_gas_fireplace,
in.misc_gas_grill,
in.misc_gas_lighting,
in.misc_hot_tub_spa,
in.misc_pool,
in.misc_pool_heater,
in.misc_pool_pump,
in.misc_well_pump,
in.occupants,
in.orientation,
in.plug_load_diversity,
in.puma,
in.puma_metro_status,
in.pv_orientation,
in.pv_system_size,
in.range_spot_vent_hour,
in.reeds_balancing_area,
in.refrigerator,
in.roof_material,
in.state,
in.tenure,
in.usage_level,
in.vacancy_status,
in.vintage,
in.vintage_acs,
in.water_heater_efficiency,
in.water_heater_fuel,
in.weather_file_city,
in.weather_file_latitude,
in.weather_file_longitude,
in.windows,
upgrade.insulation_roof,
upgrade.water_heater_efficiency,
upgrade.hvac_cooling_efficiency,
upgrade.infiltration_reduction,
upgrade.geometry_foundation_type,
upgrade.clothes_dryer,
upgrade.insulation_ceiling,
upgrade.ducts,
upgrade.hvac_heating_type,
upgrade.insulation_wall,
upgrade.insulation_foundation_wall,
upgrade.hvac_heating_efficiency,
upgrade.cooking_range,
out.electricity.ceiling_fan.energy_consumption,
out.electricity.clothes_dryer.energy_consumption,
out.electricity.clothes_washer.energy_consumption,
out.electricity.cooling_fans_pumps.energy_consumption,
out.electricity.cooling.energy_consumption,
out.electricity.dishwasher.energy_consumption,
out.electricity.freezer.energy_consumption,
out.electricity.heating_fans_pumps.energy_consumption,
out.electricity.heating.energy_consumption,
out.electricity.hot_tub_heater.energy_consumption,
out.electricity.hot_tub_pump.energy_consumption,
out.electricity.hot_water.energy_consumption,
out.electricity.lighting_exterior.energy_consumption,
out.electricity.lighting_garage.energy_consumption,
out.electricity.lighting_interior.energy_consumption,
out.electricity.mech_vent.energy_consumption,
out.electricity.plug_loads.energy_consumption,
out.electricity.pool_heater.energy_consumption,
out.electricity.pool_pump.energy_consumption,
out.electricity.pv.energy_consumption,
out.electricity.range_oven.energy_consumption,
out.electricity.refrigerator.energy_consumption,
out.electricity.well_pump.energy_consumption,
out.natural_gas.fireplace.energy_consumption,
out.natural_gas.grill.energy_consumption,
out.natural_gas.hot_tub_heater.energy_consumption,
out.natural_gas.lighting.energy_consumption,
out.natural_gas.pool_heater.energy_consumption,
time,
total_energy_consumption
  )

# Separate character and numeric variables
character_vars <- select_if(predictor_variables, is.character)

numeric_vars <- select_if(predictor_variables, is.numeric)


# One-hot encode character variables
encoded_character <- character_vars %>%
  mutate_all(as.factor) %>%
  mutate_all(funs(as.numeric(as.factor(.))))

# Combine the numeric and encoded character variables
predictor_variables <- bind_cols(numeric_vars, encoded_character)

predictor_variables
# Calculate correlations with missing value handling
correlations <- sapply(predictor_variables, function(var) cor(var, data$total_energy_consumption, use = "pairwise.complete.obs"))

# Extract the absolute correlation values
correlation_with_energy_consumption <- abs(correlations)

# Sort the features by their absolute correlation values
sorted_features <- sort(correlation_with_energy_consumption, decreasing = TRUE)

# Print the sorted features and their correlation values
result <- data.frame(Feature = names(sorted_features), Correlation = sorted_features)
view(result)



```
b. Feature Importance with Machine Learning Models: Train a machine learning model (e.g., Random Forest, Gradient Boosting) on your data and evaluate feature importance. Many machine learning libraries provide built-in functions to calculate feature importance.

```{r}
library(randomForest)

# Fit a Random Forest model
rf_model <- randomForest(total_energy_consumption ~ ., data = predictor_variables)

# Get feature importance scores
importance_scores <- importance(rf_model)

# Sort features by importance
sorted_features <- importance_scores[order(importance_scores, decreasing = TRUE), , drop = FALSE]

# Print the sorted features and their importance scores
view(sorted_features)



```


c. Recursive Feature Elimination (RFE): Use RFE with a machine learning model to iteratively remove less important features and see how it impacts model performance.

```{r}
library(caret)

# Create a control function for RFE
ctrl <- rfeControl(functions = rfFuncs, method = "cv", number = 10)

# Run RFE with a Random Forest model
rfe_result <- rfe(predictor_variables, data$total_energy_consumption, sizes = 1:ncol(predictor_variables), rfeControl = ctrl)

# Print the results
print(rfe_result)


```




```{r}



```
```{r}

```





```{r}
df <- read_csv('C:/Users/nehac/Downloads/Project_mock.csv')
energy_consumption_columns <- data %>%
  select(matches("^out\\..*consumption$"))

# Calculate the total energy consumption for each row
data$total_energy_consumption <- rowSums(energy_consumption_columns, na.rm = TRUE)

```



```{r}

# Load required libraries
library(dplyr)
library(tidyr)

# Select relevant predictor variables (features) and the target variable (energy consumption)
predictor_variables <- data %>%
  select(bldg_id,
upgrade,
weight,
applicability,
in.sqft,
in.bedrooms,
in.building_america_climate_zone,
in.ceiling_fan,
in.city,
in.clothes_dryer,
in.clothes_washer,
in.clothes_washer_presence,
in.cooking_range,
in.cooling_setpoint,
in.corridor,
in.county,
in.dishwasher,
in.federal_poverty_level,
in.geometry_floor_area_bin,
in.geometry_foundation_type,
in.geometry_garage,
in.geometry_stories,
in.geometry_stories_low_rise,
in.heating_fuel,
in.heating_setpoint,
in.heating_setpoint_has_offset,
in.heating_setpoint_offset_magnitude,
in.hot_water_fixtures,
in.hvac_cooling_efficiency,
in.hvac_cooling_partial_space_conditioning,
in.hvac_cooling_type,
in.hvac_has_ducts,
in.hvac_has_zonal_electric_heating,
in.hvac_heating_efficiency,
in.hvac_heating_type,
in.hvac_heating_type_and_fuel,
in.income,
in.income_recs_2015,
in.income_recs_2020,
in.infiltration,
in.insulation_ceiling,
in.insulation_floor,
in.insulation_foundation_wall,
in.insulation_rim_joist,
in.insulation_roof,
in.insulation_slab,
in.insulation_wall,
in.lighting,
in.misc_extra_refrigerator,
in.misc_freezer,
in.misc_gas_fireplace,
in.misc_gas_grill,
in.misc_gas_lighting,
in.misc_hot_tub_spa,
in.misc_pool,
in.misc_pool_heater,
in.misc_pool_pump,
in.misc_well_pump,
in.occupants,
in.orientation,
in.plug_load_diversity,
in.puma,
in.puma_metro_status,
in.pv_orientation,
in.pv_system_size,
in.range_spot_vent_hour,
in.reeds_balancing_area,
in.refrigerator,
in.roof_material,
in.state,
in.tenure,
in.usage_level,
in.vacancy_status,
in.vintage,
in.vintage_acs,
in.water_heater_efficiency,
in.water_heater_fuel,
in.weather_file_city,
in.weather_file_latitude,
in.weather_file_longitude,
in.windows,
upgrade.insulation_roof,
upgrade.water_heater_efficiency,
upgrade.hvac_cooling_efficiency,
upgrade.infiltration_reduction,
upgrade.geometry_foundation_type,
upgrade.clothes_dryer,
upgrade.insulation_ceiling,
upgrade.ducts,
upgrade.hvac_heating_type,
upgrade.insulation_wall,
upgrade.insulation_foundation_wall,
upgrade.hvac_heating_efficiency,
upgrade.cooking_range,
out.electricity.ceiling_fan.energy_consumption,
out.electricity.clothes_dryer.energy_consumption,
out.electricity.clothes_washer.energy_consumption,
out.electricity.cooling_fans_pumps.energy_consumption,
out.electricity.cooling.energy_consumption,
out.electricity.dishwasher.energy_consumption,
out.electricity.freezer.energy_consumption,
out.electricity.heating_fans_pumps.energy_consumption,
out.electricity.heating.energy_consumption,
out.electricity.hot_tub_heater.energy_consumption,
out.electricity.hot_tub_pump.energy_consumption,
out.electricity.hot_water.energy_consumption,
out.electricity.lighting_exterior.energy_consumption,
out.electricity.lighting_garage.energy_consumption,
out.electricity.lighting_interior.energy_consumption,
out.electricity.mech_vent.energy_consumption,
out.electricity.plug_loads.energy_consumption,
out.electricity.pool_heater.energy_consumption,
out.electricity.pool_pump.energy_consumption,
out.electricity.pv.energy_consumption,
out.electricity.range_oven.energy_consumption,
out.electricity.refrigerator.energy_consumption,
out.electricity.well_pump.energy_consumption,
out.natural_gas.fireplace.energy_consumption,
out.natural_gas.grill.energy_consumption,
out.natural_gas.hot_tub_heater.energy_consumption,
out.natural_gas.lighting.energy_consumption,
out.natural_gas.pool_heater.energy_consumption,
time,
total_energy_consumption
  )

# Separate character and numeric variables
character_vars <- select_if(predictor_variables, is.character)
numeric_vars <- select_if(predictor_variables, is.numeric)

# One-hot encode character variables
encoded_character <- character_vars %>%
  mutate_all(as.factor) %>%
  mutate_all(funs(as.numeric(as.factor(.))))

# Combine the numeric and encoded character variables
predictor_variables <- bind_cols(numeric_vars, encoded_character)

# Calculate correlation coefficients
correlations <- cor(predictor_variables, use = "complete.obs", method = "pearson")

# Extract the absolute correlation values for the target variable
correlation_with_energy_consumption <- abs(correlations[, "total_energy_consumption"])

# Sort the features by their absolute correlation values
sorted_features <- sort(correlation_with_energy_consumption, decreasing = TRUE)

# Print the sorted features and their correlation values
result <- data.frame(Feature = names(sorted_features), Correlation = sorted_features)
print(result)





```


```{r}
#in.sqft - Square footage of the area, with a correlation coefficient of about 0.311.
#in.occupants - Number of occupants, with a correlation of approximately 0.190.
# Heating setpoint, which might be the temperature set on a thermostat for heating, with a correlation of around 0.176.
#in.bedrooms - Number of bedrooms, with a correlation coefficient of about 0.142.
#in.cooling_setpoint - Cooling setpoint, likely similar to heating setpoint but for cooling, with a correlation of around 0.131.
#The other features listed (in.insulation_floor, in.insulation_wall, in.heating_fuel, in.insulation_ceiling, in.windows, in.insulation_foundation_wall) have even smaller correlation coefficients, suggesting they have a less significant linear relationship with total energy consumption in this specific dataset.
```




```{r}
#__________________________fresh Start__________________________________




library(ggplot2)
library(dplyr)
library(lubridate)  # For date manipulation


# Convert the "time" column to a DateTime object
data$time <- as.POSIXct(data$time, format = "%d-%m-%Y %H:%M")

# Aggregate energy consumption by day
daily_energy_consumption <- data %>%
  group_by(Date = as.Date(floor_date(time, unit = "days"))) %>%
  summarize(DailyEnergyConsumption = sum(total_energy_consumption, na.rm = TRUE))

# Create a ggplot object for the daily energy consumption trend
ggplot(daily_energy_consumption, aes(x = Date, y = DailyEnergyConsumption)) +
  geom_line() +  # Create a line chart
  labs(
    x = "Date",
    y = "Daily Energy Consumption",
    title = "Daily Energy Consumption Trend"
  ) +
  theme_minimal()


```
NEW EDA

```{r}
library(readr)
data <- read_csv("merged_df.csv")
head(data)

```

1.What is the county wise energy consumption? (Jash n Sampada)
```{r}
ggplot(data, aes(x = in.county, y = totalconsumption, fill = in.county)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Energy Consumption per County",
       x = "County",
       y = "Total Energy Consumption") + theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

2.What is the distribution of energy usage in July across residential properties in South Carolina and the relevant parts of North Carolina? (Neha,Jash,Sampada)
```{r}
library(tidyverse)
# Filter the data for properties in South Carolina and relevant parts of North Carolina
filtered_data <- data %>%
  filter(in.state %in% c("SC", "NC"))  # Replace "NC" with relevant parts of North Carolina

# Convert 'in.sqft' to numeric if it's not already
filtered_data$in.sqft <- as.numeric(filtered_data$in.sqft)

# Filter the data for the month of July
july_data <- filtered_data

# Create a factor variable for in.usage_level
july_data$in.usage_level <- factor(july_data$in.usage_level)

# Create a custom color palette
custom_colors <- c("green", "orange", "red")

# Create a histogram to visualize the distribution of energy usage in July
# Use the 'fill' aesthetic to specify custom colors
ggplot(july_data, aes(x = in.usage_level, fill = in.usage_level)) +
  geom_bar() +
  scale_fill_manual(values = custom_colors) +  # Specify custom colors
  labs(
    title = "Distribution of Energy Usage in July",
    x = "Energy Usage Level",
    y = "Frequency"
  )
```
3.How does energy usage in July vary across different counties within South Carolina ? (Neha,Sampada,Jash)

```{r}
filtered_data <- data %>% filter(in.state %in% c("SC", "NC"))
# Calculate the total energy consumption by county
county_energy_consumption <- filtered_data %>%
  group_by(in.county) %>%
  summarize(TotalEnergyConsumption = sum(totalconsumption, na.rm = TRUE)) %>%
  arrange(desc(TotalEnergyConsumption))  # Sort by TotalEnergyConsumption in descending order

# Create a ggplot object to visualize county-wise energy consumption
ggplot(county_energy_consumption, aes(x = TotalEnergyConsumption, y = fct_reorder(in.county, desc(TotalEnergyConsumption)))) +
  geom_bar(stat = "identity", fill = "blue") +  # Create a horizontal bar plot
  labs(
    x = "Total Energy Consumption",
    y = "County",
    title = "County-wise Energy Consumption (Descending Order) for South Carilona"
  ) +
  theme_minimal()
```


```{r}
library(readr)
data2 <- read_csv("merged_geo_avgwd_df.csv")
head(data2)
```
4. How does temperature effect the energy consumption in North and South Carilona (Jash n sampada)

```{r}
filtered_data <- data2 %>% filter(in.state %in% c("SC", "NC"))
ggplot(filtered_data, aes(x = avg_temp, y = totalconsumption)) +
  geom_point(size = 3, color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Temperature vs. Energy Consumption",
       x = "Temperature (Celsius)",
       y = "Energy Consumption") + theme_minimal() #Has a positive correlation
```

5. How does humidity effect the energy consumption in North and South Carilona (Jash n sampada)

```{r}
filtered_data <- data2 %>% filter(in.state %in% c("SC", "NC"))
ggplot(filtered_data, aes(x = avgHumidity, y = totalconsumption)) +
  geom_point(size = 3, color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Humidity vs. Energy Consumption",
       x = "Humidity", y = "Energy Consumption") + theme_minimal() #Has a negative correlation
```

6. How does income effect the energy consumption in North and South Carilona (Jash n sampada)

```{r}
# Assuming 'data' is your data frame
# Replace 'in.income' and 'totalconsumption' with your actual column names

custom_colors <- c("green", "orange", "red")

ggplot(data, aes(x = in.income, fill = totalconsumption)) +
  geom_bar(position = "stack", stat = "count") +
  scale_fill_manual(values = custom_colors) +
  labs(title = "Income vs Usage Levels",
       x = "Income Range",
       y = "Power consumption") + theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```



```{r}

```

7. Daily Energy Consumption (Neha)

```{r}
library(ggplot2)
library(dplyr)
library(lubridate)  # For date manipulation


# Convert the "time" column to a DateTime object
data$time <- as.POSIXct(data$time, format = "%d-%m-%Y %H:%M")

# Aggregate energy consumption by day
daily_energy_consumption <- data %>%
  group_by(Date = as.Date(floor_date(time, unit = "days"))) %>%
  summarize(DailyEnergyConsumption = sum(totalconsumption, na.rm = TRUE))

# Create a ggplot object for the daily energy consumption trend
ggplot(daily_energy_consumption, aes(x = Date, y = DailyEnergyConsumption)) +
  geom_line() +  # Create a line chart
  labs(
    x = "Date",
    y = "Daily Energy Consumption",
    title = "Daily Energy Consumption Trend"
  ) + theme_minimal()
```

8. Energy consumption for 6 appliances

```{r}
# Load the ggplot2 library
library(ggplot2)
filtered_data <- data2 %>% filter(in.state %in% c('SC', 'NC'))
# Sample data (replace this with your actual data)
energy_data2 <- data.frame(
  EnergyConsumed = c(filtered_data$sum_ceiling_fan, filtered_data$sum_clothes_dryer,filtered_data$sum_freezer,filtered_data$sum_cooling_fans_pumps,filtered_data$sum_dishwasher,filtered_data$sum_heating_fans_pumps),Appliance = c('fan','dryer','freezer','cooling fan pumps','dishwasher','heating fan pumps' )
)

# Create a pie chart
ggplot(energy_data2, aes(x = "", y = EnergyConsumed, fill = Appliance)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  labs(title = "Energy Consumption by Appliance",
       fill = "Appliance",
       y = "Energy Consumed (kWh)") +
  scale_fill_manual(values = c("fan" = "orange", "dryer" = "blue", "freezer" = "green","cooling fan pumps" = 'yellow',"dishwasher" = 'red' , "heating fan pumps" = "pink")) +
```

9. Are there any noticeable outliers in energy consumption that could indicate data entry errors or exceptional usage scenarios? (Neha)

```{r}
# Select the energy consumption column (replace with the appropriate column name)
energy_consumption <- data$totalconsumption

# Create a box plot to visualize the distribution of energy consumption
ggplot(data, aes(x=in.county,y = energy_consumption)) +
  geom_boxplot(fill = "lightblue", color = "blue") +
  labs(
    y = "Energy Consumption",
    title = "Box Plot of Energy Consumption"
  ) +
  theme_minimal() +  coord_flip()

# Calculate summary statistics to identify outliers
summary_stats <- summary(energy_consumption)
Q1 <- quantile(energy_consumption, 0.25)
Q3 <- quantile(energy_consumption, 0.75)
IQR <- Q3 - Q1
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR

# Find outliers
outliers <- data %>%
  filter(energy_consumption < lower_bound | energy_consumption > upper_bound)

# Print summary statistics and outliers
cat("Summary Statistics:\n")
cat(paste("Min:", min(energy_consumption), "\n"))
cat(paste("1st Quartile (Q1):", Q1, "\n"))
cat(paste("Median (Q2):", median(energy_consumption), "\n"))
cat(paste("3rd Quartile (Q3):", Q3, "\n"))
cat(paste("Max:", max(energy_consumption), "\n"))
cat(paste("IQR (Interquartile Range):", IQR, "\n"))
cat(paste("Lower Bound:", lower_bound, "\n"))
cat(paste("Upper Bound:", upper_bound, "\n"))

cat("\nOutliers:\n")
print(outliers)
```

10. The buildings in the above counties that has maximum power consumption. (Neha)
```{r}
# Load required libraries
library(dplyr)


# Filter data for specific counties
counties_to_search <- c("G4500190", "G4500790", "G4500130", "G4500510", "G4500910", "G4500350", "G4500450")

filtered_data <- data %>%
  filter(in.county %in% counties_to_search)

# Find the building with maximum power consumption in each county
max_power_buildings <- filtered_data %>%
  group_by(in.county) %>%
  top_n(1, totalconsumption) %>%
  ungroup()

# Print the result
head(max_power_buildings)
```

11. Does vintage houses have higher energy consumption (Jash n sampada)

```{r}
# Create a bar plot
ggplot(data, aes(x = in.vintage, y = totalconsumption, fill = in.vintage)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(title = "Relationship between Vintage Houses and Total Energy Consumption",
       x = "Vintage",
       y = "Total Energy Consumption",
       fill = "Vintage") + theme_minimal()
```

12. Percentage Distribution of Usage Levels by Climate Zone and City (Jash and Neha)

```{r}
ggplot(data, aes(x = in.city, fill = in.usage_level)) +
  geom_bar(position = "fill") +
  facet_wrap(~in.building_america_climate_zone) +
  labs(title = "Percentage Distribution of Usage Levels by Climate Zone and City",
       x = "City",
       y = "Percentage",
       fill = "Usage Level") +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

 13. Appliances vs Energy Consumption chart.(Sai and Jash)
```{r}

library(ggplot2)
filtered_data <- merged_geo_avgwd_df %>% filter(in.state %in% c('SC', 'NC'))
# Sample data (replace this with your actual data)
energy_data2 <- data.frame(
  EnergyConsumed = c(filtered_data$sum_ceiling_fan, filtered_data$sum_clothes_dryer,filtered_data$sum_freezer,filtered_data$sum_lighting_exterior,filtered_data$sum_dishwasher,filtered_data$sum_lighting_gas,filtered_data$sum_range_oven,filtered_data$sum_lighting_interior,filtered_data$sum_hot_water,filtered_data$sum_hot_tub_heater,filtered_data$sum_cooling),Appliance = c('fan','dryer','freezer','lighting exterior','dishwasher','lighting gas','range oven','lighting interior','hot water','hot tub heater','cooling' )
)

# Create a pie chart
ggplot(energy_data2, aes(x = Appliance, y = EnergyConsumed, fill = Appliance)) +
  geom_bar(stat = "identity", width = 1, position = 'stack') +
  labs(title = "Energy Consumption by Appliance",
       fill = "Appliance",
       y = "Energy Consumed (kWh)")
```
```{r}
unique(merged_df$in.clothes_dryer)
```


```{r}
out_column_names
```



```



